Social Visual Analytics: DevOps for Data Scientists

### 2015-01-31 system.tex
yes

\stephen{Let's work from front to back: from human interface to the
services and frameworks that support it.}

@@

Figure~\ref{fig:something} shows the RCloud developer interface.
The center panel is a notebook for code and visualizations.
Code is edited in this panel, and visualizations are rendered.
Supplementary ``asset'' files, which are part of the notebook,
though not part of its executable flow, are edited on the right.
Controls at the top of the screen allow the user to run, view, and
share the notebook. The call mechanism is the URL itself; arguments are
read from the URL and the final cell is the result. On the left is a
browsing area for searching and seeing other users' notebooks, and a
help system.

@@

The main unit of computation in RCloud is a \emph{notebook}.
A notebook holds a sequence of \emph{cells}, each of which contains a
snippet of code or hypertext in Markdown. This is not a novel idea;
executable documents structured this way are present in many
other systems, including Mathematica, IPython and Sage.

One of the main contributions of RCloud is the notion that notebooks
are ``always deployed''. This improves the ability to share and modify
experiments and compose the results without binding to specific versions
of a notebook.
On the other hand, there are situations where it might be important to
be able to name a specific version. We do not expect designers to 
decide which versions need to be preserved, but we embrace
\emph{transparent} versioning.
This is similar to models like Jankun-Kelly et al.'s p-set calculus \cite{Jankun-Kelly:2007:MFV}
and VisTrails's version tree \cite{Callahan:2006:VVM}, where every change in the state of the system is tracked.

To implement this, it is fortunate that we can build RCloud on top of
Github's \emph{gists}~\cite{Github:2014:GG}. Github offers a HTTP interface
for creating simplified git repositories, the main limitation being the
restriction to text-only files in a single directory. The GitHub web-service
API provides most of the semantics we need for the versioning
portion of the storage back end: access to previous versions,
comments, starring, and forking.

Using GitHub for storage and versioning also has the advantage of allowing
us to use other Github capabilities that can be invoked by HTTP and JSON.
Particularly, we can easily provide full search, using Apache SOLR
\stephen{whatever that is!}
%
This perspective has implications for
systems-based research and prototyping in visual analytics:
for small teams to efficiently explore novel ideas toward the
improvement of a large, complex software ecosystem, it pays
to use existing solutions as far as possible.
%
Even though it involved learning standards that are not directly
related to visual analytics, our early decision to use the GitHub
gist API was successful. Future projects can gain many
of the same benefits by adopting a similar strategy.

@@

In RCloud, reputation and interest are a relationship between
\emph{notebooks} and \emph{users}, rather than a relationship between
user pairs. We chose this approach because we expect initial 
RCloud deployments to have relatively few users, but some users will
relatively many notebooks. Under that assumption, assigning
interest to users would not provide sufficiently ``high-resolution'' data.

We incorporate both explicit and implicit indications of interest
in notebooks. Explicit interest is signaled by ``starring,'' or
clicking on a button that marks a notebook as interesting.
This makes explicit indication of interest a nearly trivial operation,
always readily available, to encourage its use.

Implicit signaling of interest is supported by keeping click-through
\cite{Joachims:2005:AIC} and execution counts. (In addition to these
classic techniques of collecting feedback from web search, we anticipate
applying static and dynamic code analysis to infer fine-grained
information about relationships, for example, which packages and data
sets often appear together.)

@@

Because notebooks are continually published, any work is immediately
available both as a subroutine and a visual component. Close colleagues
can start on the next stage of analysis, or delve into the data,
even while the original author is polishing an algorithm or its
presentation. The code is the page, and can either be shaped into
a function of inputs and outputs, or the linear cells of the notebook
can be reworked into a full-fledged HTML layout.

>>>>

\stephen{Let's work from front to back, from human interface to the
services and frameworks that support it.}

@@

Figure~\ref{fig:something} shows the RCloud developer interface.
The center panel is a notebook for code and visualizations.
Code is edited in this panel, and visualizations are rendered.
An inventory of supplementary ``asset'' files, that are part of
the notebook, though not in its executable flow, are edited on the right.
Controls at the top of the screen allow the user to run, view, and
share the notebook. The result of code execution is the final cell of
the notebook.  On the left is a browsing area for searching and viewing
other users' notebooks, and a help system.

The call mechanism is the URL itself; arguments are read from the URL
and the final cell is rendered as the result.  \stephen{That needs
more explanation.}

@@

The main unit of computation in RCloud is a \emph{notebook}.
A notebook holds a sequence of \emph{cells}, each of which contains a
snippet of code or hypertext in Markdown. This is not a novel idea;
executable documents structured this way are a feature of many
other systems, including Mathematica, IPython and Sage.
\stephen{Code is executed when...?}

One of the main contributions of RCloud is the idea that notebooks
are ``always deployed''. In other words, the most recent version of
a notebook is immediately available.\stephen{What if it is broken?}
This makes it convenient to share and modify experiments and compose
results without binding to a specific version of a notebook.
On the other hand, there are situations where it might be important
to name specific versions. We do not expect designers to decide which
versions need to be preserved, but we embrace \emph{transparent} versioning.
This is similar to models like Jankun-Kelly et al.'s p-set calculus \cite{Jankun-Kelly:2007:MFV}
and VisTrails's version tree \cite{Callahan:2006:VVM}, where every change in the state of the system is tracked.

To implement this, we built RCloud on top of Github's \emph{gists}~\cite{Github:2014:GG}.
Github offers a HTTP interface for creating simplified git repositories, the main limitation
being a restriction to text-only files in a single directory. The GitHub web-service
API provides most of the semantics we need for the versioning portion of the storage back end:
access to previous versions, comments, starring, and forking.

Using GitHub for storage and versioning also exposes other capabilities
that can be invoked with JSON and HTTP.
Particularly, we can provide full text search using Apache SOLR.
SOLR is scalable, can index in near-realtime, supports multiple character sets,
indexes several common types of documents, and has schemas and faceted search.
%
Integrating existing services and software components, especially open source,
instead of implementing custom code is a trend in the software industry, and
is very positive for systems research and prototyping in visual analytics.
By adopting existing technology, small teams with limited resoures can explore
novel ideas toward the improvement of large, complex software ecosystems.
In our case, even though this approach involved learning standards not directly
related to visual analytics, our early decision to use the GitHub gist API
turned out to be successful. Future projects could gain many of the same benefits
by adopting this strategy. In fact, it will be essential for technology adoption
and transfer, because it is almost impossible for any one tool or platform to
``own'' or support the entire visual analytics process.
\stephen{Are we preaching too much?}

@@

Information retrieval based on collecting usage and recommendations
is a cornerstone of modern web services. We would like to help data
scientists to find workbooks (and therefore items in their contents
such as code, data, and colleagues with specific expertise) with
the benefit of such information.

In RCloud, reputation and interest are a relationship between
\emph{notebooks} and \emph{users}, rather than a relationship between
user pairs. We chose this approach because we expect initial 
RCloud deployments to have relatively few users, but some users to
create many notebooks. Under that assumption, assigning interest
to users would not provide sufficiently ``high-resolution'' data.

We incorporate both explicit and implicit indications of interest
in notebooks. Explicit interest is indicated by ``starring,'' or
clicking on a button that marks a notebook as interesting.
This makes explicit indication of interest a nearly trivial operation,
always readily available, to encourage its use.

Implicit signaling of interest is supported by keeping click-through counts
\cite{Joachims:2005:AIC} and execution counts. (In addition to these
standard techniques of collecting feedback from web search, we anticipate
applying static and dynamic code analysis to infer fine-grained
information about relationships, for example, which packages and data
sets often appear together.)

@@

Because notebooks are continually published, any work is immediately
available both as a subroutine and a visual component. Close colleagues
can start on the next stage of analysis, or delve into the data,
even while the original author is polishing an algorithm or its
presentation. \stephen{We glossed over the things that can go wrong.
We should at least discuss that briefly.}
The code is the page, and can either be shaped into
a function of inputs and outputs, or the linear cells of the notebook
can be reworked into a full-fledged HTML layout.
